---
title: "p8105_hw6_ver2117_A"
author: "Varvy Rousseau"
date: "2022-12-03"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Loading relevant packages & settings
```{r}
library(tidyverse)
library(modelr)
library(tidyr)
library(p8105.datasets)
library(dplyr)
knitr::opts_chunk$set(
	  echo = TRUE,
	  warning = FALSE,
	  fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
  )
  
theme_set(theme_minimal() + theme(legend.position = "bottom"))
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis")
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

#Problem 2
#Loading in Homicide Dataset
```{r}
homi_df = read.csv("data/homicide-data1.csv", na = c("", "NA", "Unknown")) 
```

#Create a city_state variable (e.g. “Baltimore, MD”), and a binary variable indicating whether the homicide is solved. Omit cities Dallas, TX; Phoenix, AZ; and Kansas City, MO – these don’t report victim race. Also omit Tulsa, AL – this is a data entry mistake. For this problem, limit your analysis those for whom victim_race is white or black. Be sure that victim_age is numeric.

```{r}
homi_df = homi_df %>%
  mutate(
    city_state = paste(city, state, sep = ",")) %>%
    select(-city, -state) %>%
  mutate(
    victim_age = as.numeric(victim_age),
    disposition = sapply(disposition, switch,
                       "Closed without arrest" = 0,
                       "Open/No arrest" = 0,
                       "Closed by arrest" = 1)) %>%
  relocate(city_state, .after = victim_sex) %>%
  filter(city_state != "Tulsa,AL") %>%
  filter(victim_race %in% c("Black", "White")) %>%
  select(city_state, disposition, victim_race, victim_age, victim_sex)

```
#For the city of Baltimore, MD, use the glm function to fit a logistic regression with resolved vs unresolved as the outcome and victim age, sex and race as predictors. Save the output of glm as an R object; apply the broom::tidy to this object; and obtain the estimate and confidence interval of the adjusted odds ratio for solving homicides comparing male victims to female victims keeping all other variables fixed.

```{r}
Bmore = homi_df %>%
  filter(city_state == "Baltimore,MD")

Bmo_log = glm(disposition ~ victim_age + victim_race + victim_sex,
      data = Bmore, family = binomial())%>%
  broom::tidy() %>% 
  mutate(OR = exp(estimate), 
         CI_higher = exp(estimate +1.96 * std.error),
         CI_lower = exp(estimate -1.96 * std.error)) %>%
  select(term, log_OR = estimate, OR, p.value, starts_with("CI")) %>% 
  knitr::kable(digits = 3)

Bmo_log

```

#Now run glm for each of the cities in your dataset, and extract the adjusted odds ratio (and CI) for solving homicides comparing male victims to female victims. Do this within a “tidy” pipeline, making use of purrr::map, list columns, and unnest as necessary to create a dataframe with estimated ORs and CIs for each city.
```{r}
city_log = 
  homi_df %>%
  nest(data = -city_state) %>%
  mutate(
    models = map(.x = data, ~ glm(disposition ~ victim_age + victim_sex + victim_race, data = .x, family = binomial())), results = map(models, broom::tidy)) %>%
  select(results, city_state) %>%
  unnest(results) %>%
  mutate(OR = exp(estimate), 
         CI_higher = exp(estimate +1.96 * std.error),
         CI_lower = exp(estimate -1.96 * std.error)) %>%
  select(term, log_OR = estimate, OR, p.value, starts_with("CI"))

```


#Create a plot that shows the estimated ORs and CIs for each city. Organize cities according to estimated OR, and comment on the plot.

```{r}
city_log %>%
  filter(term == "victim_sexMale") %>%
  mutate(
    city_state = fct_reorder(city_state, OR)) %>%
  ggplot(aes(x = city_state, y = OR)) + geom_point() +geom_errorbar(aes(ymin = CI_lower, ymax = CI_higher)) + theme(axis.text.x = element_text(angle = 80, hjust = 1)) + labs(title = "Plot of Estimated ORs and CIs for Each City", x = "City State", y = "Odds Ratio")
 
```
Looking at this plot is appears that, Albuquerque,NM has the highest upper whisker and upper quartile of odds ratio while Chicago, IL has the lowest.  It is also interesting how the medians of ORs have been relatively consistent throughout each citystate. 


##Problem 3
#Loading the Birthweight Dataset
```{r}
birth_df = read_csv("data/birthweight.csv")
birth_df
```

## Cleaning the birthweight dataset
```{r}
birth_w = 
  birth_df %>%
  janitor::clean_names() %>%
  mutate(
    frace = as.factor(frace),
    mrace = as.factor(mrace),
    babysex = as.factor(babysex),
    malform = as.factor(malform)) 

sapply(birth_w, function(x) sum(is.na(x)))

```
After loading and tidying the total number of observations is `r nrow(birth_w)`. There are `r nrow(birth_w)` rows and `r ncol(birth_w)` columns. Important variables include: babysex, bhead,
blength, bwt, delwt, fincome,frace, gaweeks, malform, menarche, mheigth, momage, mrace, parity, pnumlbw, pnumgsa, ppbmi, ppwt, smoken, wtgain. 



### My regression model
For my regression model I selected these variables as well as the others ones listed have a higher effect on a child's birthweight. The variables that I have chosen are:
- mrace: Mother's race
- fincome: Family monthly income
- gaweeks: Gestational age in weeks
- parity:Number of live births prior to this pregnancy
- wtgain: Mother's weight gain during pregnancy
- bwt: Baby's birth weight

According to Kelli Komro and her colleagues in the article:"Effects of State-Level Earned Income Tax Credit Laws on Birth Outcomes by Race and Ethnicity," there has been a difference in factors that effect birth weight.  Factors that had a difference among races were the gestational weeks, the family's income, and the mother's race which led me to include those variables in my model.  I also wanted to include the variable parity and the mother's weight gain during pregnancy to see if those variables also made an impact on birth weight. The link to the article is: https://www.ncbi.nlm.nih.gov/pmc/articles/PMC6419088/#:~:text=Birth%20weight%20gains%20are%20higher,with%20white%20mothers%2C%20except%20among

```{r}
fit_1 = lm(bwt ~ mrace + fincome + gaweeks + parity + wtgain, data = birth_w)
fit_1
```

```{r}
fit_1 %>%
  broom::glance()

fit_1 %>%
  broom::tidy()

fit_1 %>% 
  broom::tidy() %>% 
  select(term, estimate, p.value) %>% 
  knitr::kable(digits = 3)
```

#Plot of Modoel Residuals Against Fitted Values
```{r}
birth_w %>%
  modelr::add_predictions(fit_1) %>%
  modelr::add_residuals(fit_1) %>%
  ggplot(aes(x = pred, y = resid)) + geom_point(alpha = 0.4) + geom_smooth(color = "blue", method = "lm", se = FALSE) + labs(
    title = " Plot of Model Residuals Against Predicted Values",
    x = "Predicted Values",
    y = "Residual Values") 

```

# Compare your model to two others: (1)One using length at birth and gestational age as predictors (main effects only);(2)One using head circumference, length, sex, and all interactions (including the three-way interaction) between these. We are comparing these  models using use crossv_mc and functions in purrr as appropriate.

```{r}
mod_1 = lm(bwt ~ blength + gaweeks, data = birth_w)
mod_2 = lm(bwt ~ bhead + babysex + blength + bhead*blength + bhead*babysex + blength*babysex + bhead*babysex*blength, data = birth_w)

cv_df = crossv_mc(birth_w, 100) %>%
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble))
  

cv_df = 
  cv_df %>%
  mutate(
   BW_model = map(train, ~lm(bwt ~ mrace + fincome + gaweeks + parity + wtgain, data = .x)),
   Model_O = map(train, ~lm(bwt ~ blength + gaweeks, data = .x)),
   Model_T = map(train, ~lm(bwt ~ bhead + babysex + blength + bhead*blength + bhead*babysex + blength*babysex + bhead*babysex*blength, data = .x))) %>%
   mutate(
    rmse_bw = map2_dbl(BW_model, test, ~rmse(model = .x, data = .y)),
    rmse_mO = map2_dbl(Model_O, test, ~rmse(model = .x, data = .y)),
    rmse_mT = map2_dbl(Model_T, test, ~rmse(model = .x, data = .y)))
 

```  
  
  
Here is both a Boxplot and a Violin plots of the prediction error distribution:

#Boxplot
```{r}
cv_df %>% 
  select(starts_with("rmse")) %>% 
  pivot_longer(
    everything(),
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_") %>% 
  mutate(model = fct_inorder(model)) %>% 
  ggplot(aes(x = model, y = rmse)) + geom_boxplot() + labs(title = "RMSE Distributions Across of the Models", x = "Model", y = "RMSE")
```

#Violin plot
```{r}
cv_df %>% 
  select(starts_with("rmse")) %>% 
  pivot_longer(
    everything(),
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_") %>% 
  mutate(model = fct_inorder(model)) %>% 
  ggplot(aes(x = model, y = rmse)) + geom_violin() + labs(title = "RMSE Distributions Across of the Models", x = "Model", y = "RMSE")
```
The plots that I made composing of all three models have shown that Model Two is the best fit because it has the lowest RMSE compared to the other two models. Model Two is composed of the head circumference, length, sex, and all interactions.  It seems as if my Model that I made was not the best as it had the highest RMSE out of all three. :(
